SIR Model With Testing Flows and Isolation

### 2020-11-09 codes/ali_test_sir.Rmd
some insights updated

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.4  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_t <- seq(0,1, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])

@@

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
beta  <- 0.35  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.1,1.5 , length.out=n_in) #note omega must be > rho, was 0.011

eta_t <- seq(0.001,0.01, length.out=n_out)  
s <- seq(1,2,length.out=n_out)

 df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)

@@

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

@@

**Interpretation of the R0 contour plot**

Note that $\eta_t\leq \eta_w$, i.e., reported people have a lower transmission probability than awaiting individuals. The perfect isolation/quarantine occurs when $\eta_w=\eta_t=0$ (see the upper left corner panel). Thus, we only consider the lower-triangle panels in the plot [not sure how to extract them in ggplot?].
(1) Looking at the first column of panels, there seems to be a compensatory relation between $\rho$ and $\omega$. i.e., higher $\rho$ requires smaller $\omega$ to keep R0 lower than 1. Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $\rho$ to achieve the critical R0.
(2) A sensible result that we are expecting is $\partial{R_0}/\partial{\omega}<0$.

>>>>

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_t <- seq(0,1, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
 W_S <- .3 #choose 0.3 for TTI testing plan and 1 for the random testing.

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])

@@

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case 
rho     <- seq(0,0.01, length.out=n_in)
omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
eta_t <- seq(0.001,0.1, length.out=n_out)
s <- seq(1,10,length.out=n_out)

 df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)

@@

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
That is, the lower isolation the higher $R_0$. Note that perfect isolation occurs at $\eta_w = 0$. 
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

@@

**Insights from the R0-contour plots**

1. TTI requires lower testing rate than random testing to achieve $R_0=1$.

2. When random testing regime is applied, $W_S=W_I=W_R$,

(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
 - variations of the $\eta_w$ has a negligible effect on the critical testing.
 - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates. This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$. Maybe because more people will move to $I_u$ [?, not very sure here.]
 (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true.

3. When TTI regime is applied,
